---
title: "Practical Machine Learning Course Project"
author: "Mykola Sydorov"
date: "13 02 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## About

The goal of the project is to predict the manner of performing unilateral dumbbell biceps curls based on data from accelerometers of 6 participants.

### Data Loading
```
```{r}
library(caret)
library(rattle)
library(randomForest)
urlTrain <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
urlTest <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"

TrData<-read.csv(urlTrain, header = T, na.strings = c("", "NA"))
TeData <- read.csv(urlTest, header = T, na.strings = c("", "NA"))

TrData$classe <- as.factor(TrData$classe) 
head(TrData)
```

## Removing variables that are mostly NA
```{r}
NAtr <- apply(TrData,2,function(x) {sum(is.na(x))}) 
TrData <- TrData[,which(NAtr == 0)]
NAte <- apply(TeData,2,function(x) {sum(is.na(x))}) 
TeData <- TeData[,which(NAte == 0)]
```

## Removing nonnumerical variables
```{r}
nmrk <- which(lapply(TrData, class) %in% "numeric")
preObj <-preProcess(TrData[,nmrk],method=c('knnImpute', 'center', 'scale'))
TrDataN <- predict(preObj, TrData[,nmrk])

TeDataN <-predict(preObj,TeData[,nmrk])
TrDataN$classe <- TrData$classe
```

## Removing variables with nearly zero variance
```{r}
nzv <- nearZeroVar(TrDataN,saveMetrics=TRUE)
TrDataN <- TrDataN[, nzv$nzv==FALSE]

nzv <- nearZeroVar(TeDataN,saveMetrics=TRUE)
TeDataN <- TeDataN[, nzv$nzv==FALSE]
```

##Cross Validation set
```{r}
set.seed(0703)
inTrain = createDataPartition(TrDataN$classe, p = .75, list=FALSE)
training = TrDataN[inTrain,]
testing = TrDataN[-inTrain,]
```

## Classification tree
```{r}
modfit_tree <- train(classe ~ ., method="rpart", data=training)
fancyRpartPlot(modfit_tree$finalModel)
```

#Trainig model
```{r}
mtry <- tuneRF(training[,-28], training$classe, ntreeTry=500, stepFactor=1.5,improve=0.01, 
               plot=FALSE, trace=TRUE, dobest=FALSE)
modfit_rf <- randomForest(classe~.,data=training, mtry=9, ntree=500)

```

##Testing the model on dataset
```{r}
pred <- predict(modfit_rf, testing)
confusionMatrix(pred, testing$classe)
```

##Prediction
```{r}
pred_val <- predict(modfit_rf, TeDataN)
pred_val
```